---
title: "PS270 Final Project"
author: "Josie Tucker"
description: "My final project, 4/29/26"
output:
  distill::distill_article:
    self_contained: false
---



```{r echo = FALSE}
library(readr)
library(tidyverse)
library(knitr)
library(broom)
```

```{r echo = FALSE}
my_data <- read_csv("data/qog_bas_cs_jan26 (1).csv")
```

 
```{r echo = FALSE}
plotting_data <- my_data |>
select(cname, wdi_expedu, vdem_gender) |>
  na.omit()
```




## Introduction

**Research Question:** Does a countries government spending on education correlate with higher levels of female political empowerment?

**Hypothesis:** My null hypothesis is that there IS NO statistically significant positive relationship between government spending on education and the women's political empowerment index. My alternative hypothesis is that there IS a statistically significant positive relationship between government spending on education and the women's political empowerment index. 

I predict that the data and results will show a statistically significant positive relationship between government spending on education and the women's political empowerment index scores. 

**Significance:** This research question holds significance as it provides insight to a factor that potentially enables or holds back women from feeling empowered politically in their countries. In efforts to close the gender gap in politics and international relations, focusing on potential driving factors like education allow for structural, policy based changes that will lead to real impact. Furthermore, this topic is just generally interesting to see if there is a correlation between these two factors, and if having more well-funded education systems leads to women feeling more empowered to participate in politics. 


## Data

This data is sourced from the University of Gothenberg's Quality of Government (GoQ) Basic Dataset (2026) on 156 different countries. The **predictor variable** is the countries Government Spending on Education which is indicated by the wdi_expedu variable. This variable is calculated by the countries total public expenditure on education as a percentage of the countries GDP. The **outcome variable** is the Women's Political Empowerment Index, indicated by the . This variable is measured on a 0-1 scale, and depicts female involvement in a variety of political spheres, activities, and institutions within the country. 

The **criteria for rejecting the null hypothesis** is if the p-value is less than or equal to the alpha level of 0.05, and/or the coefficient is negative. 
The **criteria for failing to reject the null hypothesis** is if the p-value is greater than the alpha level of 0.05, and the coefficient is positive. 

This histogram plot shows the distribution of the Women's Political Empowerment Scores on the Index score of 0-1.  


```{r echo = FALSE}
distribution <- ggplot(data = plotting_data) +
  aes(x = vdem_gender) +
  geom_histogram(binwidth = 0.05, fill = "steelblue", color = "white") +
  labs(
    x = "Women's Political Empowerment Index (0-1)",
    y = "Number of Countries",
    title = "Distribution of Women's Political Empowerment Scores",
    caption = "Source: University of Gothenburg QoG Basic Dataset (2026)"
  ) +
  theme_minimal()
distribution
```


## Results



```{r echo = FALSE}
tibble_data <- my_data |>
  select(cname, wdi_expedu, vdem_gender) |>
  na.omit()

country_table <- tibble_data |>
  head(15) |>
  knitr::kable(
    digits = 3,
    col.names = c("Country", "Education Spending (% of GDP)", "Empowerment Index")  )

country_table

```
(Table showing 15 out of 156 countries)




**Scatter Plot of Two Variables Relationship:**

```{r echo = FALSE}
  ggplot(data = plotting_data) +
  aes(x = wdi_expedu, y = vdem_gender) +
  geom_point() +
   labs(x = "Government Spending on Education",
       y = "Womens Political Empowerment Index",
       title = "Government Spending and Womens Political Empowerment",
       source = "University of Gothenburg, 2026")
```
**Scatter Plot with Regression:**


```{r echo = FALSE}
  ggplot(data = plotting_data) +
  aes(x = wdi_expedu, y = vdem_gender) +
  geom_point() + geom_smooth(method = "lm") +
   labs(x = "Government Spending on Education",
       y = "Womens Political Empowerment Index",
       title = "Gov. Spending and Womens Political Empowerment With Regression",
       source = "University of Gothenburg, 2026")
```


**Interpretation**: 

The above plots and tables show the distribution of the University of Gothenberg's 2026 data set between a countries government spending on education and their women's political empowerment index. The second scatter plot includes a further visual analysis, with a regression line included, demonstrating the positive relationship between the two variables with the blue line. 







**Statistical Regression:**



```{r echo = FALSE}
model <- lm(vdem_gender ~ wdi_expedu, data = plotting_data)
summary(model)
```




```{r echo = FALSE}
tidy(model) |> 
  knitr::kable(
    digits = 3, 
    col.names = c("Term", "Estimate", "Std. Error", "t-statistic", "p-value")
  )
```



**Interpretation and Evaluation**:

These results demonstrate a statistically significant positive relationship between a countries spending on education and the women's empowerment index of the country. The **p-value of 0.00006127** (rounded down to 0) is much smaller than the alpha level of 0.05, meaning that the likelihood this positive relationship between the two variables happened by chance is very small. 

Furthermore, the **positive coefficient of 0.0287** indicates that as a countries spending on education increases by 1 unit, we would expect their women's empowerment index to correspondingly increase by approximately 0.03 units. While this might seem small, 0.03 units on a scale of 0-1 could have real-world policy related implications regarding spending on education. 

Finally, the **multiple R-squared variable of 0.099** indicates that spending on education accounts for around 10% of the variation in countries women's empowerment scores. So while education spending does play a role in the scores, there are other factors also contributing to them that should be accounted for. 


These results allow us to conclude that there is strong statistical evidence of a positive correlation between government spending on education and that countries women's political empowerment index. However, this relationship is **not to be interpreted casually**, because this linear regression can not conclude that spending exclusively creates higher empowerment, only that the two variables are associated/correlated. 



## Conclusion

**Summary of Results**:

These distributions and regression analysis show strong statistical support for rejecting the null hypothesis and supporting the alternative hypothesis that there is a statistically significant positive relationship between government education spending and the Women’s Political Empowerment Index. 


**Assessment of Extent for Support for Hypothesis**:

With a positive coefficient of 0.0287 and a p-value of 0.00006127 (well below the alpha level of 0.05) the data suggests that increased investment in education correlates with higher levels of female political empowerment, and allows us to reject the null hypothesis and support the alternative with high confidence. 


**Limitations of the Analysis and Threats to Inference**:

There are limitations to these results, and they should be interpreted solely as correlated and not casual. Possible confounding variables include factors such as a countries overall wealth or regime type (level of democratization). Furthermore, the lower r squared value of 0.099 means that education spending accounts for only approximately 10% of the variation in the women's political empowerment scores, meaning a lot of other factors account for 90% of the remaining variation. An example of these factors could be the social norms of a country, meaning how supported/allowed women are to be involved in politics or education. With more time and resources, this analysis could be improved by including multiple regressions that incorporate some of these control variables like regime type or gender equality levels to help isolate the impact of education spending. Finally, this data set from the University of Gothenburg only accounts for 156 countries, including the missing countries would add further information to these findings. All of these listed things should be considered when analyzing the real-world effects of government spending on women's political empowerment and potential policy changes.



**AI Disclosure Statement**:

I utilized artificial intelligence for a few different aspects of my final project. Firstly, for some of my more difficult/complex code chucks I had to partly rely on AI to produce the correct outcomes. Secondly, some error messages or issues knitting were occurring that I could not figure out how to solve. I used AI to help resolve these issues and successfully push the correct parts of my project to the website. Finally, I used AI to check over my written analysis and responses and make sure they were correctly interpreting my results. 



**Code Printout**: 

```{r print-code, ref.label=knitr::all_labels(), echo = TRUE, eval=FALSE}

```

