---
title: "PS270 Final Project"
author: "Josie Tucker"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---




## Milestone 1 - Project thoughts

I am interested in exploring data related to a social issue I care about or policy surrounding it. 


## Milestone 2 - Data and Proposal 
For my final project, I will be using the University of Gothenburg’s Quality of Government (QoG) Basic Dataset. My research question is whether government spending on education leads to higher levels of female political empowerment. I hypothesize that countries with higher education spending levels will lead to higher levels of female political empowerment. I theorize this because when a state prioritizes and invests in education, that leads to citizens, especially women, educated and passionate enough to get involved in politics and make their voices heard. Better schools and education shows a commitment to overall social development and improvement, including the empowerment of women in the political spheres. My two variables of interest are government spending on education (wdi_expedu variable in dataset) and the women's political empowerment index (Vdem_gender in dataset). I will compare countries with high education spendings womens political empowerment indexes to those with medium and low spending to compare. Which countries I will include/not include in my studies will be decided later on in my project timeline. If spending fluctuations align with empowerment indexes I will know my hypothesis is correct. For instance, high spending = high empowerment index, medium spending = medium empowerment index, and low spending = low empowerment index. If there is no clear relationship between these two factors then I know my hypothesis is disproved. 


```{r}
library(readr)
library(tidyverse)
```

```{r}
my_data <- read_csv("data/qog_bas_cs_jan26 (1).csv")
```



## Milestone 3 - One data visualization (organized table and scatter plot)


```{r}
plotting_data <- my_data |>
select(cname, wdi_expedu, vdem_gender) |>
  na.omit()
plotting_data
```


```{r}
  ggplot(data = plotting_data) +
  aes(x = wdi_expedu, y = vdem_gender) +
  geom_point() +
   labs(x = "Government Spending on Education",
       y = "Womens Political Empowerment Index",
       title = "Government Spending and Womens Political Empowerment",
       source = "University of Gothenburg, 2026")
```

## Milestone 4 - Analysis of Results
The above tibble and scatter plot from milestone 3 show the distribution of the University of Gothenberg's 2026 data set between a countries government spending on education and their womens political empowerment index. The scatter plot below includes a second visual analysis, this time with a regression line, indicating the relationship between the two variables.  

The scatter plot shows that as a countries spending on education increases, so does there womens political empowerment index. 


```{r}
  ggplot(data = plotting_data) +
  aes(x = wdi_expedu, y = vdem_gender) +
  geom_point() + geom_smooth(method = "lm") +
   labs(x = "Government Spending on Education",
       y = "Womens Political Empowerment Index",
       title = "Government Spending and Womens Political Empowerment",
       source = "University of Gothenburg, 2026")
```
**Statistical Regression Analysis:**


```{r}
model <- lm(vdem_gender ~ wdi_expedu, data = plotting_data)
summary(model)
```

These results demonstrate a statistically significant positive relationship between a countries spending on education and the womens empowerment index of the country. The p-value of 0.00006127 is much smaller than an alpha level of 0.05, meaning that the likelihood this positive relationship between the two variables happened by chance is very small. 

Furthermore, the coefficient of 0.0287 indicates that as a countries spending on education increases by 1 unit, we would expect their women's empowerment index to correspondingly increase by approximately 0.03 units. 

Finally, the multiple R-squared variable, of 0.09 indicates that spending on education accounts for around 10% of the variation in countries womens empowerment scores. So while education spending does play a role in the scores, there are other factors also contributing to them that should be accounted for. 


## Final Report 

**Introduction**

Research Question: Does a countries government spending on education correlate with higher levels of female political empowerment?

Hypothesis: My null hypothesis is that there IS NO statistically significant positive relationship between government spending on education and the women's political empowerment index. My alternative hypothesis is that there IS a statistically significant positive relationship between government spending on education and the women's political empowerment index. 

I predict?

Significance: This research question holds significance as it provides insight to a factor that potentially enables or holds back women from feeling empowered politically in their countries. In efforts to close the gender gap in politics and international relations, focusing on potential driving factors like education allow for structural, policy based changes that will lead to real impact. Furthermore, this topic is just generally interesting to see if there is a correlation between these two factors, and if having more well-funded education systems leads to women feeling more empowered to participate in politics. 


**Data**

This data is sourced from the University of Gothenberg's Quality of Government (GoQ) Basic Dataset (2026) on 156 different countries. The **predictor variable** is the countries Government Spending on Education which is indicated by the wdi_expedu variable. This variable is calculated by the countries total public expenditure on education as a percentage of the countries GDP. The **outcome variable** is the Women's Political Empowerment Index. This variable is measured on a 0-1 scale, and depicts female involvement in a variety of political spheres, activities, and institutions within the country. 

The criteria for rejecting the null hypothesis is if the p-value is less than or equal to the alpha level of 0.05, and/or the coefficient is negative. 
The criteria for failing to reject the null hypothesis is if the p-value is greater than the alpha level of 0.05, and the coefficient is positive. 

This histogram plot shows the distribution of the Women's Political Empowerment Scores on the Index score of 0-1.  


```{r, echo=false}
ggplot(data = plotting_data) +
  aes(x = vdem_gender) +
  geom_histogram(binwidth = 0.05, fill = "steelblue", color = "white") +
  labs(
    x = "Women's Political Empowerment Index (0-1)",
    y = "Number of Countries",
    title = "Distribution of Women's Political Empowerment Scores",
    caption = "Source: University of Gothenburg QoG Basic Dataset (2026)"
  ) +
  theme_minimal()
```






```{r print-code, ref.label=knitr::all_labels(), echo = TRUE, echo=FALSE}

```


